Sentence Generation as Planning with Probabilistic LTAG\\

### 2010-05-21 trunk/Papers/pcrisp-10/experiments.tex
Minor changes. Camera ready. :)

We estimate our probability models on section 1 to 23 of the converted WSJ using maximum likelihood estimation. We use section 0 as a testing set. However, since the number of \pcrisp\ operators grows quadratically with the grammar size, generating long sentences requires too much time to run batch experiments.  We therefore restricted our evaluation to the 416 sentences in Section 0 that are shorter than 16 words.

For this testing set we automatically create semantic representation for each sentence, by instantiating the semantic content of each elementary used in its derivation. We use these representations as input for our system and compare the system output against the original sentence.

@@

Table \ref{results} presents the results of the experiment for five different generation systems.  We compare three variants of \pcrisp: the fully lexicalized PTAG model (``PTAG''), the fully unlexicalized model (``unlexicalized''), and a linear interpolation model in which we (manually) set the weight of level 2 to 0.9 \todo{stimmt das?} and the weight of level 3 to 0 (``interpolation'').  We also list results for the non-statistical \crisp\ system of  \newcite{kollerstone2007} (``\crisp'') and the greedy search heuristic used by \spud.  All these systems are based on a reimplementation of the FF planner \cite{hoffmannnebel2001}, to which we added a search heuristic that takes action costs into account for the \pcrisp\ systems.

>>>>

We estimate our probability models on section 1 to 23 of the converted WSJ using maximum likelihood estimation. We use section 0 as a testing set. However, since the number of \pcrisp\ operators grows quadratically with the grammar size, generating long sentences requires too much time to run batch experiments.  We therefore restrict our evaluation to the 416 sentences in Section 0 that are shorter than 16 words.

For this testing set we automatically create semantic representation for each sentence, by instantiating the semantic content of each elementary tree used in its derivation. We use these representations as input for our system and compare the system output against the original sentence.

@@

Table \ref{results} presents the results of the experiment for five different generation systems.  We compare three variants of \pcrisp: the fully lexicalized PTAG model (``PTAG''), the fully unlexicalized model (``unlexicalized''), and a linear interpolation model (``interpolation'') in which we (manually) set the weight of level 2 to 0.9 and the weight of level 3 to 0.  We also list results for the non-statistical \crisp\ system of  \newcite{kollerstone2007} (``\crisp'') and the greedy search heuristic used by \spud.  All these systems are based on a reimplementation of the FF planner \cite{hoffmannnebel2001}, to which we added a search heuristic that takes action costs into account for the \pcrisp\ systems.

### 2010-05-21 trunk/Papers/pcrisp-10/introduction.tex
Minor changes. Camera ready. :)

Many sentence generation systems are organized in a pipeline
architecture, in which the input semantic representation is first
enriched, e.g.\ by referring expressions, by a \emph{sentence
  planner} and only then transformed into natural language strings by
a \emph{surface realizer} \cite{reiterdale2000}.  An alternative
approach is \emph{integrated} sentence generation, in which both steps
are performed by the same algorithm, as in the SPUD system
\cite{Stone2003a}.  An integrated algorithm can sometimes generate
better and more succinct sentences \cite{stone98textual}.  SPUD itself
gives up some of this advantage by using a greedy search heuristic
for efficiency reasons.  The CRISP system, a recent reimplementation
of SPUD using search techniques from AI planning, achieves high
efficiency without sacrificing complete search
\cite{kollerstone2007,KolHof10}.

>>>>

Many sentence generation systems are organized in a pipeline
architecture, in which the input semantic representation is first
enriched, e.g.\ with referring expressions, by a \emph{sentence
  planner} and only then transformed into natural language strings by
a \emph{surface realizer} \cite{reiterdale2000}.  An alternative
approach is \emph{integrated} sentence generation, in which both steps
are performed by the same algorithm, as in the SPUD system
\cite{Stone2003a}.  An integrated algorithm can sometimes generate
better and more succinct sentences \cite{stone98textual}.  SPUD itself
gives up some of this advantage by using a greedy search heuristic
for efficiency reasons.  The CRISP system, a recent reimplementation
of SPUD using search techniques from AI planning, achieves high
efficiency without sacrificing complete search
\cite{kollerstone2007,KolHof10}.

### 2010-05-21 trunk/Papers/pcrisp-10/pcrisp.tex
Minor changes. Camera ready. :)

While we leave the basic sentence generation mechanism intact, we need to modify the concrete formulation of \crisp\ planning operators to accommodate bilexical dependencies. Likewise, we need to take the step from classical planning to \emph{metric} planning systems which can use the probabilities.
In metric planning \cite{fox2002}, planning actions can modify the value of numeric variables in addition to adding and deleting logical literals to the state. The goal state specifies constraints on this variable. In the simplest case the variable can only be increased by a static cost value in each action, and the goal state contains the objective to minimize the total cost. While systems such as Metric-{\sc FF} \cite{hoffmann2003} do not guarantee optimality, they do generally offer good results. We address our encoding of sentence generation with {\sc ptag} as metric planning in section \ref{ssec:pcrisp-domains}.

>>>>

While we leave the basic sentence generation mechanism intact, we need to modify the concrete formulation of \crisp\ planning operators to accommodate bilexical dependencies. Likewise, we need to take the step from classical planning to \emph{metric} planning systems which can use the probabilities.
In metric planning \cite{fox2002}, planning actions can modify the value of numeric variables in addition to adding and deleting logical literals from the state. The goal state specifies constraints on this variable. In the simplest case the variable can only be increased by a static cost value in each action, and the goal state contains the objective to minimize the total cost. While systems such as Metric-{\sc FF} \cite{hoffmann2003} do not guarantee optimality, they do generally offer good results. We address our encoding of sentence generation with {\sc ptag} as metric planning in section \ref{ssec:pcrisp-domains}.

### 2010-05-21 trunk/Papers/pcrisp-10/related.tex
Minor changes. Camera ready. :)

Statistical methods are popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their {\sc nitrogen} system. Such systems generate a set of candidate sentences using a (possibly overgenerating) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as {\sc HAL}ogen \cite{langkildeknight1998,langkilde2000} and OpenCCG \cite{whitebaldridge2003}.  The {\sc Fergus} system \cite{bangalorerambow2000} is a variant of this approach which, like {\sc pcrisp}, employs {\sc ltag}. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using using only these trees with an $n$-gram model on the output sentence. An alternative to the generate and select approach is to use a probabilistic grammar model, like {\sc   ptag}, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both models can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar in that it uses {\sc ptag} to find the most likely output structure. However, the previous work discussed so far addresses surface realization only. We extend this to a statistical NLG algorithm which does surface realization and sentence planning at the same time.

Our treatment of integrated sentence planning and surface realization as planning, is inherited directly from {\sc crisp} \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms \cite{hoffmannnebel2001} spawned a renewed interest in planning for NLG.  {\sc Crisp} uses such algorithms to efficiently solve the sentence generation problem defined by {\sc spud} \cite{Stone2003a}. {\sc Spud}, which instead used an incomplete greedy algorithm, is based on an {\sc ltag} whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the {\sc spud} problem realizes this goal and simultaneously selects referring expressions. The next section explores {\sc crisp} in more detail.

>>>>

Statistical methods are popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their {\sc nitrogen} system. Such systems generate a set of candidate sentences using a (possibly overgenerating) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as {\sc HAL}ogen \cite{langkildeknight1998,langkilde2000} and OpenCCG \cite{whitebaldridge2003}.  The {\sc Fergus} system \cite{bangalorerambow2000} is a variant of this approach which, like {\sc pcrisp}, employs {\sc ltag}. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using only these trees with an $n$-gram model on the output sentence. An alternative to the $n$-gram based generate and select approach is to use a probabilistic grammar model, like {\sc   ptag}, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both models can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar in that it uses {\sc ptag} to find the most likely output structure. However, the previous work discussed so far addresses surface realization only. We extend this to a statistical NLG algorithm which does surface realization and sentence planning at the same time.

Our treatment of integrated sentence planning and surface realization as planning is inherited directly from {\sc crisp} \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms \cite{hoffmannnebel2001} spawned a renewed interest in planning for NLG.  {\sc Crisp} uses such algorithms to efficiently solve the sentence generation problem defined by {\sc spud} \cite{Stone2003a}. {\sc Spud}, which instead uses an incomplete greedy algorithm, is based on an {\sc ltag} whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the {\sc spud} problem realizes this goal and simultaneously selects referring expressions. The next section explores {\sc crisp} in more detail.
